Rapid technological progress in white light-emitting diodes \\ and its sources in innovation and technology spillovers

### 2023-03-24 2_SSL_EES/article/article_supplementary.tex
updated figure, table and equation naming
updated bibliography

\renewcommand{\figurename}{Figure SI}
\renewcommand{\tablename}{Table SI}

@@

\section{Additional Performance Metrics}

@@

Efficacy in lighting is dependent on the luminosity function, which describes the wavelength-dependent sensitivity of the human eye. A light source emitting very \textit{efficiently} in the infrared yet emitting no visible light has a very low \textit{efficacy}. The luminous efficacy of radiation $K$ is mathematically defined as the normalized, integrated product of the spectral radiant flux of a light source with the wavelength dependent human sensitivity to light \cite{cie-term-effrad}

@@

This metric can be computed from spectral data alone and does not require additional spectral normalization. It enables straightforward comparison between the performance of different downconversion phosphors, as shown in  \cref{fig:se_efficacy} in chapter \cref{ch:results}. Light sources emitting in the far red or blue part of the spectrum have lower efficacy of radiation, as the product with the luminosity function of  \cref{fig:tristimulus} shows. Care must be taken not to confuse this efficacy metric with efficacy of source, which depends not only on the spectral power distribution of a light source.

This metric describes the match of a light-emitting diode package spectrum to the human visual system.

@@

In contrast to the efficacy \textit{of radiation} $K$, the luminous efficacy \textit{of a light source} $\eta$ is defined as the ratio between the emitted luminous flux and the consumed electrical power \cite{cie-term-effsrc}

@@

Unlike the luminous efficacy of radiation, this metric cannot be extracted from the spectrum of a light source. It is often cited in device datasheets, scientific literature and textbooks when describing the performance of light-emitting diodes. Care must be taken not to confuse this efficacy metric with efficacy of radiation, which depends only on the spectral power distribution of a light source. As it captures overall device efficacy, it depends on a large number of device properties and parameters. This makes attribution to single changes in device design or manufacturing difficult.

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.95\textwidth]{./figures/cost_model.pdf}
    \caption{Schematic diagram of the cost model showing inputs to each step and computational steps leading to the cost model outputs. For the description of cost variables, see definitions for equation \cref{eqn:cost_wafer}.}
    \label{fig:costmodel-schematic}
\end{figure}

@@

Calling something here as 'upstream' and 'downstream' is kind of confusing. I suggest to replace the text highlighted by this comment with the following simplifying reformulation: 
'It is split between the two stages of the manufacturing process: first stage combining operations at the wafer level, followed by the LED packaging stage.'

A schematic diagram of the cost model developed is presented in  \cref{fig:costmodel-schematic}. The cost model is process step-based. It is split between the upstream part of the manufacturing process and the downstream part of the manufacturing process. The former is performed at the wafer level. The latter is performed at the package level. The model takes as inputs both parameters specific to manufacturing process steps (\textit{"manufacturing step parameters"}) and parameters affecting all manufacturing steps (\textit{"global parameters"}). The cost for each process step is then computed. The cost model returns returns the costs of individual manufacturing steps as well as the total manufacturing cost. It further considers the yield per step and returns the cumulative yield, the yielded cost per step and the yielded total manufacturing cost.

@@

\cref{fig:manuf_vtf_2012-1}-\cref{fig:manuf_csp_2020-2} show a simplified rendering of the manufacturing process of three different chip architectures considered in the cost model.

@@

    \begin{landscape}
        \begin{figure}
            \includegraphics[width=555pt]{./figures/classical_overview_2003.pdf}
            \caption{Manufacturing process for a classical LED package with lateral current spreading, circa 2003. Abbreviations: MQW - multiple quantum well; PLLC - plastic leaded chip carrier. Continued on next page.}
            \label{fig:manuf_classical_2003}
        \end{figure}
    \end{landscape}

@@

    \begin{landscape}
        \begin{figure}
            \includegraphics[width=595pt]{./figures/csp_overview_2020-2.pdf}
            \caption{(2/2) Continued from previous page. Abbreviations: Encaps. - encapsulation.}
            \label{fig:manuf_csp_2020-2}
        \end{figure}
    \end{landscape}

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=5.5cm]{./figures/die-per-wafer.pdf}
    \caption{Number of die per wafer for the different wafer sized used in different years: (DPW) $d(2002)=50$mm$\rightarrow851$DPW to $d(2020)=200$mm$\rightarrow26,838$DPW. Source: own elucidations, based on equation \cref{eqn:dpw}.}
    \label{fig:dpw}
\end{figure}

@@

which can be compared to equation \cref{eqn:C_2} to find a more general form:

@@

where $n$ is the total number of steps. This is due to the approximation introduced through the series approximation in equation \cref{eqn:series}.

@@

Using logarithmic differentiation, the integral from equation \cref{eqn:integral_1} can be rewritten as

@@

This approach was used to determine the effect of spillovers on device performance, displayed in  \cref{fig:breakthroughs_efficiency}. Due to time constraints, this approach could not be used to determine the effect of spillovers on manufacturing cost. Future work to quantify the impact of different cost drivers, including breakthroughs and changes in the manufacturing variables of equation \cref{eqn:cost_sum}, could this method.

@@

The complete list of sources is organized by the different technologies described in the :

\begin{table}[h!]
    \begin{tabularx}{\textwidth}{|l|X|}
    \hline
    \textbf{Technology} & \textbf{References} \\
    \hline
    LED & own research, compare \cite{zenodo_weinold_led_history} \\
    \hline
    CFL (<1984) & \cite{Bouwknegt1982}\cite{Vrenken1983} \\
    \hline
    CFL (1984-2011) & \cite{eger2018origin} \\
    \hline
    CFL (>2011) & \cite{Guan2015} \\
    \hline
    Fire, Incandescent, HID & \cite{azevedo2009transition} augmented by own calculations based on \cite{benesch1905beleuchtungswesen} \\
    \hline
    Max. efficacy & \cite{Murphy2012} (in contrast to the original value given in \cite{azevedo2009transition}) \\
    \hline
    \end{tabularx}
\end{table}

@@

The complete list of sources is organized by different sources:

\begin{table}[h!]
    \begin{tabularx}{\textwidth}{|l|X|}
    \hline
    \textbf{Source of References} & \textbf{References} \\
    \hline
    Stiftung Warentest (Germany) & \cite{Warentest2008}\cite{Warentest2009_1}\cite{Warentest2009_2}\cite{Warentest2010_1}\cite{Warentest2010_2}\cite{Warentest2011}\cite{Warentest2012}\cite{Warentest2013}\cite{Warentest2014_1}\cite{Warentest2014_2}\cite{Warentest2015}\cite{Warentest2016_1}\cite{Warentest2016_2}\cite{Warentest2018} \\
    \hline
    Konsument (Austria) & \cite{Konsument2010} \\
    \hline
    Which (UK) & \cite{Which2020} \\
    \hline
    Industry Periodicals & \cite{PM2020} \\
    \hline
    Government Reports & \cite{council2013assessment} \\
    \hline
    \end{tabularx}
\end{table}

@@

The complete list of sources is organized by different sources:

\begin{table}[h!]
    \begin{tabularx}{\textwidth}{|l|X|}
    \hline
    \textbf{Source of References} & \textbf{References} \\
    \hline
    Scientific Publications & \cite{plossl2010wafer}\cite{bierhuizen2007performance}\cite{gencc2019distributed}\cite{chong2014performance} \\
    \hline
    Patents & \href{https://worldwide.espacenet.com/patent/search?q=pn\%3DJPH11168235A}{JPH11168235A}, 
\href{https://worldwide.espacenet.com/patent/search?q=pn\%3DDE19921987A1}{DE19921987A1}, 
\href{https://worldwide.espacenet.com/patent/search?q=pn\%3DJPH11340514A}{JPH11340514A}, 
\href{https://worldwide.espacenet.com/patent/search?q=pn\%3DJP2001044498A}{JP2001044498A}, 
\href{https://worldwide.espacenet.com/patent/search?q=pn\%3DUS2005045893A1}{US2005045893A1}, 
\href{https://worldwide.espacenet.com/patent/search?q=pn\%3DUS2002093023A1}{US2002093023A1}, 
\href{https://worldwide.espacenet.com/patent/search?q=pn\%3DUS2006273339A1}{US2006273339A1}  \\
    \hline
    Other Publications & \href{https://web.archive.org/web/20170801160530/https://www.energy.gov/sites/prod/files/2015/02/f19/craford_innovation_sanfrancisco2015.pdf}{2015 presentation (www.energy.gov)} \newline
\href{https://web.archive.org/web/20170715230721/https://www.energy.gov/sites/prod/files/2016/02/f29/sun_china_raleigh2016.pdf}{2016 presentation (www.energy.gov)} \newline
\href{http://web.archive.org/web/20160425025936/https://www.slideshare.net/Yole_Developpement/yole-led-packagingjanuary2013reportsample}{2013 presentation (www.slideshare.net)} \\
    \hline
    \end{tabularx}
\end{table}

@@

The complete list of sources is organized by the sub-efficiencies (panels) described in the :

\begin{table}[h!]
    \begin{tabularx}{\textwidth}{|l|X|}
    \hline
    \textbf{Technology} & \textbf{References} \\
    \hline
    Panel A1 ($V_f$) & \cite{nichia2001data}\cite{lumi2002data}\cite{gen2005data}\cite{candlepwr2005data}\cite{lumi2006data}\cite{lumi2007data}\cite{nichia2008data}\cite{lumi2008data}\cite{osram2008data}\cite{jeong2011high}\cite{osram2012data}\cite{osram2013data}\cite{osram2014data} \newline \cite{lumi2016data_1}\cite{lumi2016data_2}\cite{epistar2017data}\cite{osram2017data_1}\cite{osram2017data_2}\cite{samsung2017data}\cite{samsung2018data}\cite{osram2018data}\cite{epistar2018data}\cite{lumi2019data} \\
    \hline
    Panel A2 (Droop) & Data calculated from luminous intensity curves of respective device datasheets: \cite{datasheet_osram_topled}\cite{osram2008data}\cite{osram2008gdplus}\cite{osram2018csp}\cite{datasheet_lumileds_lux1}\cite{lumi2008data}\cite{lumi2016data_1}\cite{lumi2016data_2}\cite{samsung2018data} \\
    \hline
    Panel B1 (IQE) & own research, compare \cite{zenodo_weinold_led_history} \newline
\cite{doe_ssl_multiyear_2006}\cite{doe_ssl_multiyear_2007}\cite{doe_ssl_multiyear_2008}\cite{doe_ssl_multiyear_2009}\cite{doe_ssl_multiyear_2010}\cite{doe_ssl_multiyear_2011}\cite{doe_ssl_multiyear_2012}\cite{doe_ssl_multiyear_2013}\cite{doe_ssl_multiyear_2014}\cite{doe_ssl_rnd_2015}\cite{doe_ssl_rnd_2016} \\
    \hline
    Panel B2 (LEE) & \cite{lee2005analysis}\cite{krames2007status}\cite{Jang2004}\cite{Horng2013}\cite{Liao2010}\cite{HungWenHuang2005}\cite{Leem2007}\cite{Huang2008}\cite{Wang2009}\cite{Huh2003}\cite{Horng2008}\cite{Gao2008}\cite{Chang2003}\cite{Zhou2012} \newline \cite{ChunJuTun2006}\cite{Hua2009}\cite{Matioli2010}
\cite{lee2005analysis}\cite{Zhu2015}\cite{Ding2015}\cite{Taki2019}\cite{Shchekin2006}\cite{Hu2016}\cite{Horng2010}\cite{Lin2016}\cite{Yue2018}\cite{Zhao2012}\cite{Zhu2015}\newline \cite{Ding2015}\cite{wierer2001high}\cite{Steigerwald2002}\cite{DaeSeobHan2006}\cite{Wang2006}\cite{Lee2007}\cite{Shen2007}\cite{Huang2006}\cite{Zhmakin2011} \\
    \hline
    \end{tabularx}
\end{table}

>>>>

\renewcommand{\thefigure}{SI\arabic{figure}}
\renewcommand{\thetable}{SI\arabic{table}}

@@

\section{Luminous Efficacy Metrics}

Figure 1 in the main publication uses luminous efficacy of radiation as the primary metric to describe progress in lighting technologies. Care must be taken not to confuse this metric with luminous efficacy of source, which is used in \ref{fig:phosphor_spectrum}.

@@

This metric describes the match of a light-emitting diode package spectrum to the human visual system. Efficacy in lighting is dependent on the luminosity function, which describes the wavelength-dependent sensitivity of the human eye. A light source emitting very \textit{efficiently} in the infrared yet emitting no visible light has a very low \textit{efficacy}. The luminous efficacy of radiation $K$ is mathematically defined as the normalized, integrated product of the spectral radiant flux of a light source with the wavelength-dependent human sensitivity to light \cite{cie-term-effrad}

@@

This metric can be computed from spectral data alone and does not require additional spectral normalization. It enables straightforward comparison between the performance of different downconversion phosphors, as shown in the top panel of \cref{fig:phosphor_spectrum}. Light sources emitting in the far red or blue part of the spectrum have lower efficacy of radiation. Care must be taken not to confuse this efficacy metric with the \textit{efficacy of source} described in the following subsection.

@@

The luminous efficacy \textit{of a light source} $\eta$ is defined as the ratio between the emitted luminous flux and the consumed electrical power \cite{cie-term-effsrc}

@@

This metric is often cited in device datasheets, scientific literature and textbooks when describing the performance of light-emitting diodes. Care must be taken not to confuse this efficacy metric with the \textit{luminous efficacy of radiation}, which depends only on the spectral characteristics of a light source. As the luminous efficacy of a light source $\eta$ captures the overall device efficacy, it depends on a large number of other device properties and parameters. This makes attribution of changes in this metric to individual changes in device design or manufacturing difficult. For this reason, we do not use this metric in our study.

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.95\textwidth]{./figures/cost_model.pdf}
    \caption{Schematic diagram of the cost model showing inputs to each step and computational steps leading to the cost model outputs. For the description of cost variables, see definitions for \cref{eqn:cost_wafer}.}
    \label{fig:costmodel-schematic}
\end{figure}

@@

A schematic diagram of the cost model is presented in \cref{fig:costmodel-schematic}. The cost model is process step-based. It is split between the two stages of the manufacturing process: the first stage combining operations at the wafer level, followed by the LED packaging stage. The model takes as inputs parameters specific to individual manufacturing process steps (\textit{"manufacturing step parameters"}) and parameters affecting all manufacturing steps (\textit{"global parameters"}). The cost for each process step is then computed. The cost model returns returns the costs of individual manufacturing steps as well as the total manufacturing cost. It further considers the yield per step and returns the cumulative yield, the yielded cost per step and the yielded total manufacturing cost.

@@

\cref{fig:manuf_classical_2003}-\cref{fig:manuf_csp_2020-2} show a simplified rendering of the manufacturing process of three different chip architectures considered in the cost model.

@@

    \begin{landscape}
        \begin{figure}
            \includegraphics[width=555pt]{./figures/classical_overview_2003.pdf}
            \caption{Manufacturing process for a classical LED package with lateral current spreading, circa 2003. Abbreviations: MQW - multiple quantum well; PLLC - plastic leaded chip carrier.}
            \label{fig:manuf_classical_2003}
        \end{figure}
    \end{landscape}

@@

    \begin{landscape}
        \begin{figure}
            \includegraphics[width=595pt]{./figures/csp_overview_2020-2.pdf}
            \caption{(2/2) Manufacturing process for a chip scale package flip-chip LED chip with vertical current spreading, circa 2020. Continued from previous page. Abbreviations: Encaps. - encapsulation.}
            \label{fig:manuf_csp_2020-2}
        \end{figure}
    \end{landscape}

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=5.5cm]{./figures/die-per-wafer.pdf}
    \caption{Number of die per wafer for the different wafer sized used in different years: (DPW) $d(2002)=50$mm$\rightarrow851$DPW to $d(2020)=200$mm$\rightarrow26,838$DPW. Source: own elucidations, based on  \cref{eqn:dpw}.}
    \label{fig:dpw}
\end{figure}

@@

which can be compared to  \cref{eqn:C_2} to find a more general form:

@@

where $n$ is the total number of steps. This is due to the approximation introduced through the series approximation in  \cref{eqn:series}.

@@

Using logarithmic differentiation, the integral from  \cref{eqn:integral_1} can be rewritten as

@@

This approach was used to determine the effect of spillovers on device performance, displayed in  \cref{fig:breakthroughs_efficiency}. Due to time constraints, this approach could not be used to determine the effect of spillovers on manufacturing cost. Future work to quantify the impact of different cost drivers, including breakthroughs and changes in the manufacturing variables of  \cref{eqn:cost_sum}, could this method.

@@

The list of sources for Figure 1 in the main article is organized by different technologies:

\begin{table}[h!]
    \begin{tabularx}{\textwidth}{|l|X|}
    \hline
    \textbf{Technology} & \textbf{References} \\
    \hline
    LED & own research, compare \cite{zenodo_weinold_led_history} \\
    \hline
    CFL (<1984) & \cite{Bouwknegt1982}\cite{Vrenken1983} \\
    \hline
    CFL (1984-2011) & \cite{eger2018origin} \\
    \hline
    CFL (>2011) & \cite{Guan2015} \\
    \hline
    Fire, Incandescent, HID & \cite{azevedo2009transition} augmented by own calculations based on \cite{benesch1905beleuchtungswesen} \\
    \hline
    Max. efficacy & \cite{Murphy2012} \\
    \hline
    \end{tabularx}
\end{table}

@@

The sources for Figure 2 in the main article are grouped by the provider of data:

\begin{table}[h!]
    \begin{tabularx}{\textwidth}{|l|X|}
    \hline
    \textbf{Provider of References} & \textbf{References} \\
    \hline
    Stiftung Warentest (Germany) & \cite{Warentest2008}\cite{Warentest2009_1}\cite{Warentest2009_2}\cite{Warentest2010_1}\cite{Warentest2010_2}\cite{Warentest2011}\cite{Warentest2012}\cite{Warentest2013}\cite{Warentest2014_1}\cite{Warentest2014_2}\cite{Warentest2015}\cite{Warentest2016_1}\cite{Warentest2016_2}\cite{Warentest2018} \\
    \hline
    Konsument (Austria) & \cite{Konsument2010} \\
    \hline
    Which (UK) & \cite{Which2020} \\
    \hline
    Industry Periodicals & \cite{PM2020} \\
    \hline
    Government Reports & \cite{council2013assessment} \\
    \hline
    \end{tabularx}
\end{table}

@@

The sources for Figure 3 in the main article are grouped by their type:

\begin{table}[h!]
    \begin{tabularx}{\textwidth}{|l|X|}
    \hline
    \textbf{Type of Source} & \textbf{References} \\
    \hline
    Scientific Publications & \cite{plossl2010wafer}\cite{bierhuizen2007performance}\cite{gencc2019distributed}\cite{chong2014performance} \\
    \hline
    Patents & \href{https://worldwide.espacenet.com/patent/search?q=pn\%3DJPH11168235A}{JPH11168235A}, 
\href{https://worldwide.espacenet.com/patent/search?q=pn\%3DDE19921987A1}{DE19921987A1}, 
\href{https://worldwide.espacenet.com/patent/search?q=pn\%3DJPH11340514A}{JPH11340514A}, 
\href{https://worldwide.espacenet.com/patent/search?q=pn\%3DJP2001044498A}{JP2001044498A}, 
\href{https://worldwide.espacenet.com/patent/search?q=pn\%3DUS2005045893A1}{US2005045893A1}, 
\href{https://worldwide.espacenet.com/patent/search?q=pn\%3DUS2002093023A1}{US2002093023A1}, 
\href{https://worldwide.espacenet.com/patent/search?q=pn\%3DUS2006273339A1}{US2006273339A1}  \\
    \hline
    Other Publications & \href{https://web.archive.org/web/20170801160530/https://www.energy.gov/sites/prod/files/2015/02/f19/craford_innovation_sanfrancisco2015.pdf}{2015 presentation (www.energy.gov)} \newline
\href{https://web.archive.org/web/20170715230721/https://www.energy.gov/sites/prod/files/2016/02/f29/sun_china_raleigh2016.pdf}{2016 presentation (www.energy.gov)} \newline
\href{http://web.archive.org/web/20160425025936/https://www.slideshare.net/Yole_Developpement/yole-led-packagingjanuary2013reportsample}{2013 presentation (www.slideshare.net)} \\
    \hline
    \end{tabularx}
\end{table}

@@

The sources for Figure 4 in the main article are grouped by sub-efficiencies represented on different panels of the figure:

\begin{table}[h!]
    \begin{tabularx}{\textwidth}{|l|X|}
    \hline
    \textbf{Panel (Sub-Eff.)} & \textbf{References} \\
    \hline
    Panel A1 ($V_f$) & \cite{nichia2001data}\cite{lumi2002data}\cite{gen2005data}\cite{candlepwr2005data}\cite{lumi2006data}\cite{lumi2007data}\cite{nichia2008data}\cite{lumi2008data}\cite{osram2008data}\cite{jeong2011high}\cite{osram2012data}\cite{osram2013data}\cite{osram2014data} \newline \cite{lumi2016data_1}\cite{lumi2016data_2}\cite{epistar2017data}\cite{osram2017data_1}\cite{osram2017data_2}\cite{samsung2017data}\cite{samsung2018data}\cite{osram2018data}\cite{epistar2018data}\cite{lumi2019data} \\
    \hline
    Panel A2 (Droop) & Data calculated from luminous intensity curves of respective device datasheets: \cite{datasheet_osram_topled}\cite{osram2008data}\cite{osram2008gdplus}\cite{osram2018csp}\cite{datasheet_lumileds_lux1}\cite{lumi2008data}\cite{lumi2016data_1}\cite{lumi2016data_2}\cite{samsung2018data} \\
    \hline
    Panel B1 (IQE) & own research, compare \cite{zenodo_weinold_led_history} \newline
\cite{doe_ssl_multiyear_2006}\cite{doe_ssl_multiyear_2007}\cite{doe_ssl_multiyear_2008}\cite{doe_ssl_multiyear_2009}\cite{doe_ssl_multiyear_2010}\cite{doe_ssl_multiyear_2011}\cite{doe_ssl_multiyear_2012}\cite{doe_ssl_multiyear_2013}\cite{doe_ssl_multiyear_2014}\cite{doe_ssl_rnd_2015}\cite{doe_ssl_rnd_2016} \\
    \hline
    Panel B2 (LEE) & \cite{lee2005analysis}\cite{krames2007status}\cite{Jang2004}\cite{Horng2013}\cite{Liao2010}\cite{HungWenHuang2005}\cite{Leem2007}\cite{Huang2008}\cite{Wang2009}\cite{Huh2003}\cite{Horng2008}\cite{Gao2008}\cite{Chang2003}\cite{Zhou2012} \newline \cite{ChunJuTun2006}\cite{Hua2009}\cite{Matioli2010}
\cite{lee2005analysis}\cite{Zhu2015}\cite{Ding2015}\cite{Taki2019}\cite{Shchekin2006}\cite{Hu2016}\cite{Horng2010}\cite{Lin2016}\cite{Yue2018}\cite{Zhao2012}\cite{Zhu2015}\newline \cite{Ding2015}\cite{wierer2001high}\cite{Steigerwald2002}\cite{DaeSeobHan2006}\cite{Wang2006}\cite{Lee2007}\cite{Shen2007}\cite{Huang2006}\cite{Zhmakin2011} \\
    \hline
    \end{tabularx}
\end{table}
